Two Problems for Sophistication

### 2015-01-21 Facticity/Facticity/paper.tex
...

\begin{abstract}
Kolmogorov complexity measures the amount of information in data, but does not distinguish structure from noise. Kolmogorov's definition of the \emph{structure function} was the first attempt to measure only the structural information in data, by measuring the complexity of an objectively best model under two-part coding. Since then, many variations of this idea have been proposed, for which we use \emph{sophistication} as an umbrella term. We describe three fundamental problems with all existing proposals, showing each  to be \emph{unsound}. Consequently, we put forward the view that the problem is fundamental, it may be impossible to objectively quantify the sophistication.
\end{abstract}

\section{Introduction}
Komogorov complexity gives us a sound definition of the amount of information contained in a bitstring. There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would consider complexity. For example, a sequence of a million coin flips will almost certainly have maximal Kolmogoroc complexity, even though there is nothing complex about flipping a lot of coins---it's just hard work. For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data. This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed. In this paper, we investigate three serious problems with these definitions, and show why they are unsatisfactory. We believe that the problems are fundamental, and provide arguments why such attempts to define sophistication must fail.

The Kolmogorov complexity of a binary string $x$, denoted $K(x)$, is (roughly) the length of the shortest computer program that will output $x$. This length depends on which programming language is used, but, as shown by a core result known as the Invariance Theorem, the language affects the result only by a constant, independent of $x$. This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings. We can see it as \emph{a property of the data}.

>>>>

\begin{abstract}
Kolmogorov complexity measures the amount of information in data, but does not distinguish structure from noise. Kolmogorov's definition of the \emph{structure function} was the first attempt to measure only the structural information in data, by measuring the complexity of an objectively best model under two-part coding. Since then, many variations of this idea have been proposed, for which we use \emph{sophistication} as an umbrella term. We describe three fundamental problems with all existing proposals, showing each  to be \emph{unsound}. Consequently, we put forward the view that the problem is fundamental: it may be impossible to objectively quantify the sophistication.
\end{abstract}

\section{Introduction}
Komogorov complexity gives us a sound definition of the amount of information contained in a bitstring. There is, however, a nagging discrepancy between Kolmogorov complexity and what most people would consider complexity. For example, a sequence of a million coin flips will almost certainly have maximal Kolmogorov complexity, even though there is nothing complex about flipping a lot of coins---it's just a lot of work. For this reason, many scholars have defined alternative complexity measures in the spirit of Kolmogorov complexity, aimed at quantifying not \emph{all} information in a binary string, but only the \emph{meaningful} information in the data. This tradition started with a lecture by Kolmogorov in Talinn in 1973, but many alternative definitions of meaningful information, or \emph{sophistication},\footnotemark were proposed. In this paper, we investigate three serious problems with these definitions, and show why they are unsatisfactory. We believe that the problems are fundamental, and provide arguments why such attempts to define sophistication must fail.

The Kolmogorov complexity of a binary string $x$, is (roughly) the length of the shortest computer program that will output $x$. This length depends on which programming language is used, but, as shown by a central result, known as the Invariance Theorem\cite[Section~2.1]{li1993introduction}, the language affects the result only by a constant, independent of $x$. This means that for sufficiently complex objects, the choice of programming language becomes irrelevant and Kolmogorov complexity becomes an \emph{objective} measure of the complexity of binary strings. We can see it as \emph{a property of the data}.
